Discovery of the first quadruple gravitationally lensed quasar candidate with Pan-STARRS

### 2017-06-14 manuscript.tex
Checked Ciprian's changes

{\bf In this letter we report the serendipitous discovery of a quad lensed quasar candidate from the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS1, hereafter PS1) released images. The lensed quasar was discovered as part of a variability study of Active Galactic Nuclei selected based on Mid-Infrared characteristics \citep{sec01}, and data from the USNO Robotic Astrometric Telescope \citep[URAT;][]{zac15}. To our knowledge, if proven with spectroscopic observations, this would be the first gravitational lens discovered in the PS1 data. We note that this system is very similar to the well-studied quad lens RX J1131-1231 \citep{slu03}, in terms of image separation and overall configuration. We model the system to provide evidence in support of its lensing nature, and we conclude with the necessity of obtaining spectroscopic confirmation. Due to its position, from the ground this is not possible until the end of the year.}

PS1 is a wide-field imaging system, with a 1.8 m telescope and 7.7 deg$^2$ field of view, located on the summit of Haleakala in the Hawaiian island of Maui. The first PS1 data was release in December 2016 including both images and photometry \citep[see][]{cha16}. The 1.4 Gpixel camera consists of 60 CCDs with pixel size of 0.256 arcsec \citep{ona08, ton09}. It uses five filters (g$_{P1}$, r$_{P1}$, i$_{P1}$, z$_{P1}$, y$_{P1}$, hereafter $grizy$), similar to the ones used by the Sloan Digital Sky Survey \citep[SDSS;][]{york00}. The largest survey PS1 performs is the 3$\pi$ survey, covering the entire sky north of $-30\deg$ declination. Given the large sky coverage, resolution and depth, PS1 is expected to contain nearly 2000 gravitationally lensed quasars, of which about 300 will be quad lensed quasars \citep{ogu10}.

The structure of the paper is as follows:
In Section 2 we infer relative astrometry, photometry and photometric redshifts. In Section 3 we fit a mass model to the system. Discussions, conclusions and future works are presented in Section 4. We assume a cosmological model with $\Omega_M = 0.274$, $\Omega_L = 0.726$, and $h$ = 0.71. All magnitudes are in the AB system {\bf except for WISE, which are in the Vega system}.

\section{ASTROMETRY AND PHOTOMETRY}

@@

The simple modeling with a point spread function (PSF) constructed from stars in the PS1 field of view (FOV) left large residuals at the locations of the four objects, that were significant enough to affect the derived parameters. We suspect that this is due to spatial variation in the PSF across the PS1 FOV, and have therefore adopted a technique of fitting the system with an analytical PSF, which it is assumed will not change on the small scale represented by the system. We did this using Hostlens \citep{rus16}, a variant of glafic \citep{ogu10b} which uses $\chi^2$ minimization in order to fit point sources as well as S{\'e}rsic \citep{ser63} profiles convolved with an analytical PSF. The analytical PSF is comprised of two concentric Moffat \citep{mof69} profiles, each one with its own full-width half maximum, ellipticity, orientation, and shape parameter. The two profiles are also characterized by a relative flux of one to the other. We successfully modeled the four objects as point sources convolved with the PSF, therefore showing that they are point-like. Our strategy was to run Hostlens starting from 100 different positions in the parameter space, select the best of the resulting models, and further run 10 Markov Chain Monte Carlo (MCMC) chains using the Metropolis-Hastings algorithm around it. The chains consist of one million steps, with an acceptance rate of $\sim0.3$, and we removed the first 1/5 of these (the ``burn-in'' steps). We also checked that the chains have converged, using the method of \citet{gel95}. The MCMC-derived uncertainties between the various analytical parameters are shown in Fig.~\ref{mcmc_hostlens}, and generated using corner.py \citep{for16}. In order to quantify additional uncertainties inherent in our method due to the non-analytical nature of the ``true'' PSF, we ran 100 simulations where we add noise of similar properties to the real images, on top of the best-fit model. Here, we created the best-fit model by using a PSF constructed from nearby stars, and we ran Hostlens on each of the 100 simulations, using an analytical PSF. As a final estimate of our uncertainties, we used the maximum between the MCMC-derived uncertainties and those from the simulations. We were able to obtain much improved residuals in all bands, and we show these in Fig.~\ref{hostlens}.

@@

In addition to the PS1 data, we have also looked for archival data of this system. The system is detected in both the Gaia \citep{gaia} DR1 and PS1 catalogs; however, only A and D are deblended by the automatic pipeline. PS1 psf magnitudes for A are in general smaller than what we obtained probably due to blending, but for D they are very similar to our uncorrected magnitudes: 18.599$\pm$0.013,  18.153$\pm$0.002,  18.079$\pm$0.007,  17.843$\pm$0.0093,  17.534$\pm$0.023 in the $grizy$ bands, respectively. We use the Gaia position for component D as absolute astrometric reference position for this system: 26.792307, 46.511273. The errors for these coordinates given in the Gaia catalog are 15.6 and 8.9 mas, respectively.

@@

Finally, the system is very likely also a radio source. The NRAO VLA Sky Survey (NVSS) images show a source, 2MASX J01471020+4630433, with flux 12.6 mJy at 1.4 GHz. The listed position for this source is only $3\arcsec$ away from the center of the lens, which is comparable with NVSS astrometric errors \citep{con98}. We are currently in the process of further observing the system in the radio with the Very Large Array (VLA). Radio data is useful for the study of lensed quasars, as it provides high-resolution imaging of the source, as well as fluxes not affected by microlensing and extinction.

@@

We fit the photometry measured in Section 2 for the point-like sources and the faint red galaxy using spectral energy distribution (SED) templates. The colors of the point-like images are very similar (Table~1), consistent with being multiple images of the same source. Due to their point-like morphology, we suspect these to be quasars, and have therefore fit them with quasar templates in order to check the quality of the fit and to infer photometric redshifts. Photometric redshifts of quasars are notoriously difficult to constrain, due to the relatively featureless continua \citep{ric09} and the variable equivalent widths of the broad emission lines. Redshift was measured using two separate methods:
\begin{itemize}
\item The method of \citet{wu10} presents quasar colors as a function of redshifts. This is based on Sloan Digital Sky survey (SDSS) colors. PS1 bands are similar but not identical to the SDSS bands and we used the corrections of \citet{fin16} to obtain magnitudes in the SDSS system. Following \citet{wu10}, we minimized ${\chi}^2$ to obtain redshifts for each of the point sources. {\bf We only used three of the colors as the Y band from UKIRT is very different from the PS1 y band.} We plot $\chi^2_{\nu}$ in Figure~\ref{redshift} for each quasar image, and also by fitting all sources together. We notice that the curves are quite consistent with each other, as expected from the similar colors of the sources. In the latter case we obtain the best fit at $z = 0.820^{+0.006}_{-0.007}$. We notice a second minimum at $z \sim2.6$. {\bf The 
 $\chi^2_{\nu}$ for the minima is large, which could be due to the large spread seen in the colors of \citet{wu10}. Because of the poor fit we did not attempt to use any prior from a quasar luminosity function.}
\item LePhare \citep{arn99,ilb06} method uses quasar templates. We obtain the best fit at a redshift of $\sim 2.6-2.8$ for all four point sources.  These results agree with the second peak of the method above. The $\chi^2$ is between 13 and 125 for 4 degrees of freedom (5 filters). We find poorer fits (larger $\chi^2$) when using stellar or galactic SED templates, providing further support that the sources are quasars. In particular, the best-fit galaxy templates correspond to starburst galaxies at low redshift $\lesssim0.1$. At such low redshifts the sources would likely be resolved, not point-like. 
{\bf \item A combination of SDSS and WISE photometry provides good discriminators between high and low redshifts. \citet{pom15} uses he ratio between WISE W1 and W2 bands to the i band flux, high values are expected for quasars below z = 1. We added the fluxes for all four images in the i band and we obtain a total magnitude of 14.4. We obtain ratios W1/i = 14.4 and W2/i = 39.0, which both indicate a redshift z $>$ 2, and are not compatible with the low value of z = 0.8, which would require ratios of W1/i = 50 and W2/i = 120. We also used the diagnostics from \citet{wu12}. As seen in Fig.~13 in this paper, the colors z - W1 and W2 - W3 discriminate well between low and high redshift. We obtain z - W1 = 2.8 and W2 - W3 = 3.7, which both suggest a redshift larger than z = 2. Our low value of 0.8 would require z - W1 = 4.0 and W2 - W3 = 3.0.
}
\end{itemize}

The lens galaxy G is very faint but we managed to fit a galaxy SED with an early-type template, resulting in a best-fit $z = 0.57^{+0.20}_{-0.13}$. For this, we used the magnitudes measured in the $izy$ bands and assumed detection limits of 21 mag in the $gr$ bands, $\sim$ 2 mag fainter than the rest.

@@

URAT1 was a U. S. Naval Observatory project designed to provide accurate astrometry in the northern hemisphere down to about 18 mag. A single filter centered at $\sim$~720 nm was used, which is almost as red as the PS1 i filter. Observations at the Flagstaff Station over more than 3 years provided more than 50 observations for a large number of objects. The gravitational lens was discovered as part of project to study the variability of a sample of AGN selected based on WISE colors \citep{sec01}. We now present the results of this study for our quad lens, keeping in mind that the system was not resolved by URAT. To characterize the variability we used a typical method used for quasars, the structure function V, as presented in \citet{van04}. It uses the magnitude  differences between different epochs of observation to estimate variability at different time intervals (time lag). \citet{van04} used the structure function to characterize the variability of a large sample of Sloan quasars. We note that their study is statistical in nature and therefore only provides general properties which we will use to try to discriminate between the two photometric redshifts we obtained in the previous section.

In order to improve the results, we performed relative photometry using nearby stars in the 12 to 14 magnitude range, and removing stars that show variability and also removing some poor quality observations. We used more than 50 stars and 36 epochs over a time interval longer than two years. We estimate errors using the nearby stars with similar magnitudes. We obtained V = 0.074 and an average time lag of 297 days. According to Fig.~ 16 of \citet{van04}, this V value matches quasars at about z = 2 and therefore our photometric value of z = 2.6 is a better match than the low value of z = 0.8, which would require a V value of 0.05. We also used the absolute magnitude dependence in Fig.~11 of \citet{van04}. We estimate absolute magnitudes for our quasar using the demagnified average of images A, B and C. Assuming a composite quasar spectrum we estimate that the i-band magnitudes range between -24.8  and -25.20 at redshift 0.8 and between -27.6 and -27.9 at redshift 2.6. Again we see that the V value matches better a high luminosity quasar, and therefore the higher redshift is preferred.

{\bf Based on the variability results and the combination of WISE and PS1 colors in the previous section, we estimate the quasars redshift at about z = 2.6, which is preferred over the option of a low redshift z = 0.8.}
}

@@

For the best-fit mass model parameters, we find an Einstein radius $b = 1.932_{-0.011}^{+0.008}$, ellipticity $e_c = 0.163_{-0.070}^{+0.063}$, $e_s = -0.035_{-0.024}^{+0.023}$, and shear $\gamma_c = 0.122_{-0.016}^{+0.015}$, $\gamma_s = 0.065\pm0.005$. Here, we expressed the ellipticity and shear in quasi-cartesian coordinates, rather than polar coordinates, as the distribution of uncertainties is more gaussian. The two coordinate systems are related by $(e_c,\gamma_c) = (e,\gamma)\cos2(\theta_e,\theta_\gamma)$ and $(e_s,\gamma_s) = (e,\gamma)\sin2(\theta_e,\theta_\gamma)$. The convergence was again checked using the method of \citet{gel95}. The uncertainties were determined using MCMC as plotted in Fig.~\ref{mcmc_gravlens} with corner.py. The $\chi^2$ for this model is 0.8.

While the parameters derived above are independent of redshift, estimating the expected time delays between the images does require redshift information. Using the results of the previous Section, we have fixed the redshift of the source at 0.57. If we assume a source redshift of 0.82, we derive from the expression of the Einstein radius in a singular isothermal profile, $b=4\pi\left(\frac{\sigma}{c}\right)^2\frac{D_{ls}}{D_{s}}$, a velocity dispersion of $\sigma\sim500$ km/s. Here $c$ is the speed of light, and $D_{s},D_{ls}$ are the angular diameter distances to the source and between lens and source, respectively. This value is too large for a single galaxy, but if we employ the second source redshift estimate from Section 2, $\sim2.6$, we obtain $\sigma\sim315$ km/s, a more common value for massive early-type galaxies. We therefore assume this redshift to infer the predicted time delays, and we obtain $\Delta t_{BC} = 0.1$ days, $\Delta t_{AC} = 1.7$ days and $\Delta t_{DC} = 226.2$ days, with image C leading.

{\bf In order to check if the $izy$ magnitudes for the lensing galaxy are compatible with the redshift we used the luminosity prior from BPZ \citep{ben00} to estimate the redshift. The observed magnitudes for z=0.57 correspond to an absolute magnitude of about -23.0 to -23.5 in R. From the Faber-Jackson relation \citep{fab76} we obtain R $\sim$~-23.6 and therefore the measured magnitudes are compatible with the estimated redshift.}

@@

We would like to thank Arunav Kundu, Shobita Satyapal, Nathan Secrest, Bryan Dorland and Valeri Makarov for useful discussions and help.
The Pan-STARRS1 Surveys (PS1) and the PS1 public science archive have been made possible through contributions by the Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, the Max-Planck Society and its participating institutes, the Max Planck Institute for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, the Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las Cumbres Observatory Global Telescope Network Incorporated, the National Central University of Taiwan, the Space Telescope Science Institute, the National Aeronautics and Space Administration under Grant No. NNX08AR22G issued through the Planetary Science Division of the NASA Science Mission Directorate, the National Science Foundation Grant No. AST-1238877, the University of Maryland, Eotvos Lorand University (ELTE), the Los Alamos National Laboratory, and the Gordon and Betty Moore Foundation.
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.

@@

\bibitem[McKean et al.(2007)]{mck07} McKean J.~P., et al., 2007, MNRAS, 378, 109 
\bibitem[Moffat (1969)]{mof69} Moffat, A. F. J.\ 1969, \aap, 3, 455
\bibitem[Myers et al.(2003)]{mye03} Myers S.~T., et al., 2003, MNRAS, 341, 1

@@

\begin{figure}
\epsscale{1.0}
\plotone{f3.eps}
\caption{
Covariances between the parameters used to model the observed system in $i$-band with Hostlens. The three contour lines mark the 1-, 2- and 3-$\sigma$ limits. The vertical lines in the histograms mark the 16th, 50th and 84th percentiles. The parameters are, in order: FWHM in arcseconds, ellipticity, orientation, shape of the first Moffat profile (1,2,3,4), same for the second Moffat profile (5,6,7,8), relative flux as flux1/(flux1 + flux2) (9), sky level in counts (10), positions in pixels on the x and y axes of point-source A (11,12) and its flux in counts (13), and same for objects B,C,D (14-25).}
\label{mcmc_hostlens}
\end{figure}

>>>>

{\bf In this letter we report the serendipitous discovery of a quad lensed quasar candidate from the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS1, hereafter PS1) released images. The lensed quasar was discovered as part of a variability study of Active Galactic Nuclei selected based on Mid-Infrared characteristics \citep{sec01}, and data from the USNO Robotic Astrometric Telescope \citep[URAT;][]{zac15}. To our knowledge, if proven with spectroscopic observations, this would be the first published gravitational lens discovered in the PS1 data. We note that this system is very similar to the well-studied quad lens RX J1131-1231 \citep{slu03}, in terms of image separation and overall configuration. We model the system to provide evidence in support of its lensing nature, and we conclude with the necessity of obtaining spectroscopic confirmation. Due to its position, from the ground this is not possible until the end of the year.}

PS1 is a wide-field imaging system, with a 1.8 m telescope and 7.7 deg$^2$ field of view, located on the summit of Haleakala in the Hawaiian island of Maui. The first PS1 data was released in December 2016, including both images and catalogues \citep[see][]{cha16}. The 1.4 Gpixel camera consists of 60 CCDs with pixel size of 0.256 arcsec \citep{ona08, ton09}. It uses five filters (g$_{P1}$, r$_{P1}$, i$_{P1}$, z$_{P1}$, y$_{P1}$, hereafter $grizy$), similar to the ones used by the Sloan Digital Sky Survey \citep[SDSS;][]{york00}. The largest survey PS1 performs is the 3$\pi$ survey, covering the entire sky north of $-30\deg$ declination. Given the large sky coverage, resolution and depth, PS1 is expected to contain nearly 2000 gravitationally lensed quasars, of which about 300 will be quad lensed quasars \citep{ogu10}.

The structure of the paper is as follows:
In Section 2 we measure the relative astrometry, photometry and morphology of this system. In Section 3 we infer photometric redshifts, and in Section 4 we conduct a photometric variability study to support the photometric redshifts. In Section 5 we fit a lensing mass model to the system. We present our conclusions and future works in Section 6. We assume a cosmological model with $\Omega_M = 0.274$, $\Omega_L = 0.726$, and $h$ = 0.71. All magnitudes are in the AB system {\bf except for WISE, which are in the Vega system}.

\section{ASTROMETRY, PHOTOMETRY AND MORPHOLOGY}

@@

The simple modeling with a point spread function (PSF) constructed from stars in the PS1 field of view (FOV) left large residuals at the locations of the four objects, that were significant enough to affect the derived parameters. We suspect that this is due to spatial variation in the PSF across the PS1 FOV, and have therefore adopted a technique of fitting the system with an analytical PSF, which it is assumed will not change on the small scale represented by the system. We did this using Hostlens \citep{rus16}, a variant of glafic \citep{ogu10b} which uses $\chi^2$ minimization in order to fit point sources as well as S{\'e}rsic \citep{ser63} profiles convolved with an analytical PSF. The analytical PSF is comprised of two concentric Moffat \citep{mof69} profiles, each one with its own full-width half maximum, ellipticity, orientation, and shape parameter. The two profiles are also characterized by a relative flux of one to the other. We successfully modeled the four objects as point sources convolved with the PSF, therefore showing that they are point-like. Our strategy was to run Hostlens starting from 100 different positions in the parameter space, select the best of the resulting models, and further run 10 Markov Chain Monte Carlo (MCMC) chains using the Metropolis-Hastings algorithm around it. The chains consist of one million steps, with an acceptance rate of $\sim0.3$, and we removed the first 1/5 of these (the ``burn-in'' steps). We also checked that the chains have converged, using the method of \citet{gel95}. The MCMC-derived uncertainties between the various analytical parameters are shown in Fig.~\ref{mcmc_hostlens}, and generated using corner.py \citep{for16}. In order to quantify additional uncertainties inherent in our method due to the non-analytical nature of the ``true'' PSF, we ran 100 simulations where we added noise of similar properties to the real images, on top of the best-fit model. Here, we created the best-fit model by using a PSF constructed from nearby stars, and we ran Hostlens on each of the 100 simulations, using an analytical PSF. As a final estimate of our uncertainties, we used the maximum between the MCMC-derived uncertainties and those from the simulations. We were able to obtain much improved residuals in all bands, and we show these in Fig.~\ref{hostlens}.

@@

In addition to the PS1 data, we have also looked for archival data of this system. The system is detected in both the Gaia \citep{gaia} DR1 and PS1 catalogs; however, objects A, B and C are blended into a single object by the automatic pipeline. PS1 psf magnitudes for A are in general smaller than what we obtained probably due to blending, but for D they are very similar to our uncorrected magnitudes: 18.599$\pm$0.013,  18.153$\pm$0.002,  18.079$\pm$0.007,  17.843$\pm$0.0093,  17.534$\pm$0.023 in the $grizy$ bands, respectively. We use the Gaia position for component D as absolute astrometric reference position for this system: 26.792307, 46.511273. The errors for these coordinates given in the Gaia catalog are 15.6 and 8.9 mas, respectively.

@@

Finally, the system is very likely also a radio source. The NRAO VLA Sky Survey (NVSS) images show a source, 2MASX J01471020+4630433, with flux 12.6 mJy at 1.4 GHz. The listed position for this source is only $3\arcsec$ away from the center of the system in the optical bands, which is comparable with NVSS astrometric errors \citep{con98}. We are currently in the process of further observing the system in the radio with the Very Large Array (VLA). Radio data is useful for the study of lensed quasars, as it provides high-resolution imaging of the source, as well as fluxes not affected by microlensing and extinction.

@@

We fit the photometry measured in Section 2 for the point-like sources and the faint red galaxy using spectral energy distribution (SED) templates. The colors of the point-like images are very similar (Table~1), consistent with being multiple images of the same source. Due to their point-like morphology, we suspect these to be quasars, and have therefore fit them with quasar templates in order to check the quality of the fit and to infer photometric redshifts. Photometric redshifts of quasars are notoriously difficult to constrain, due to the relatively featureless continua \citep{ric09} and the variable equivalent widths of the broad emission lines. We measured redshifts using several methods:
\begin{itemize}
\item The method of \citet{wu10} uses derived quasar colors as a function of redshifts. This is based on Sloan Digital Sky survey (SDSS) colors. PS1 bands are similar but not identical to the SDSS bands and we used the corrections of \citet{fin16} to obtain magnitudes in the SDSS system. Following \citet{wu10}, we minimized ${\chi}^2$ to obtain redshifts for each of the point sources. {\bf We only used three of the colors as the Y band from UKIRT is very different from the PS1 y band.} We plot $\chi^2_{\nu}$ in Figure~\ref{redshift} for each quasar image, and also by fitting all sources together. We notice that the curves are quite consistent with each other, as expected from the similar colors of the sources. In the latter case we obtain the best fit at $z = 0.820^{+0.006}_{-0.007}$. We notice a second minimum at $z \sim2.6$. {\bf The $\chi^2_{\nu}$ for the minima is large, which could be due to the large spread seen in the colors of \citet{wu10}. We note that that paper does not list numerical uncertainties for the derived colors. We have also explored incorporating a prior based on the quasar luminosity function, given the un-magnified observed magnitudes of the source (see Section 5). While the prior prefers low redshift values, its effect on Fig.~\ref{redshift} is negligible.}
\item A second method uses the photometric redshift code LePhare \citep{arn99,ilb06} to fit a quasar template to the observed colors. We obtain the best fit at a redshift of $\sim 2.6-2.8$ for all four point sources. These results agree with the second peak of the method above. The $\chi^2$ is between 13 and 125 for 4 degrees of freedom (5 filters). We find poorer fits (larger $\chi^2$) when using stellar or galactic SED templates, providing further support that the sources are quasars. In particular, the best-fit galaxy templates correspond to starburst galaxies at low redshift $\lesssim0.1$. At such low redshifts the sources would likely be resolved, not point-like. 
{\bf \item A combination of SDSS and WISE photometry provides good discriminators between high and low redshift quasars. \citet{pom15} used the ratio between WISE W1 and W2 bands to the $i$ band flux, and obtained that high values are expected for quasars below redshift of 1. We added the fluxes for A, B, C, D in the $i$ band (to account for the fact that the system is not resolved in WISE; G is of negligible flux) and we obtained a total magnitude of 14.4. We obtain ratios W1/$i$ = 14.4 and W2/$i$ = 39.0, which both indicate a redshift > 2, and are not compatible with the low value of $\sim 0.8$, which would require ratios of W1/$i$ = 50 and W2/$i$ = 120. We also used the diagnostics from \citet{wu12}. As seen in Fig.~13 of that paper, the colors $z$ - W1 and W2 - W3 discriminate well between low and high redshift. We obtain $z$ - W1 = 2.8 and W2 - W3 = 3.7, which both suggest a redshift larger than 2. Our low value of 0.8 would require z - W1 = 4.0 and W2 - W3 = 3.0.
}
\end{itemize}

The lens galaxy G is very faint but we managed to fit a galaxy SED with an early-type template, resulting in a best-fit $z = 0.57^{+0.20}_{-0.13}$. For this, we used the magnitudes measured in the $izy$ bands and assumed detection limits of 21 mag in the $gr$ bands, $\sim$ 2 mag fainter than the rest. {\bf We independently checked that the observed magnitudes are reasonable for a redshift of $z=0.57$: for this redshift, the observed magnitudes imply an absolute magnitude of about $-23.0 \sim -23.5$ in $R$-band. From the Faber-Jackson relation \citep{fab76}, using the velocity dispersion we estimate from the lens modeling in Section 5, we obtain $R \sim -23.6$, therefore not inconsistent.}

@@

Here we undertake a quasar variability study as an independent check of the redshift estimate we derived in the previous section.
URAT1 was a U. S. Naval Observatory project designed to provide accurate astrometry in the northern hemisphere down to about 18 mag. A single filter centered at $\sim$~720 nm was used, which is almost as red as the PS1 $i$ filter. Observations at the Flagstaff Station over more than 3 years provided more than 50 epochs for a large number of objects. 
%The gravitational lens was discovered as part of project to study the variability of a sample of AGN selected based on WISE colors \citep{sec01}. We now present the results of this study for our quad lens, keeping in mind that the system was not resolved by URAT.

To characterize the variability of this system we used a typical method used for quasars, the structure function V, as presented in \citet{van04}. It uses the magnitude differences between different epochs of observation to estimate variability at different time intervals (time lag). \citet{van04} used the structure function to characterize the variability of a large sample of Sloan quasars. We note that their study is statistical in nature and therefore only provides general properties which we use to try to discriminate between the two photometric redshifts we obtained in the previous section.

In order to improve the results, we performed relative photometry using nearby stars in the 12 to 14 magnitude range, and removing stars that show variability and also removing some poor quality observations. We used more than 50 stars and 36 epochs spanning over two years. We estimate errors using the nearby stars with similar magnitudes. We obtained V = 0.074 and an average time lag of 297 days. According to Fig.~16 of \citet{van04}, this V value matches quasars at about $z = 2$ and therefore our photometric value of $z = 2.6$ is a better match than the low value of $z = 0.8$, which would require a V value of 0.05. We also used the absolute magnitude dependence in Fig.~11 of \citet{van04}. We estimate an absolute magnitude for our source quasar using the demagnified average of the observed magnitudes of A, B and C (see Section 5). Assuming a composite quasar spectrum we estimate that the $i$-band rest-frame absolute magnitude ranges between $-24.8$ and $-25.2$ at redshift 0.8, and between $-27.6$ and $-27.9$ at redshift 2.6. Again we see that the V value matches better a high luminosity quasar, and therefore the higher redshift is preferred. We note that in this study we have ignored variability caused by microlensing (see the next Section).}

@@

For the best-fit mass model parameters, we find an Einstein radius $b = 1.932_{-0.011}^{+0.008}$, ellipticity $e_c = 0.163_{-0.070}^{+0.063}$, $e_s = -0.035_{-0.024}^{+0.023}$, and shear $\gamma_c = 0.122_{-0.016}^{+0.015}$, $\gamma_s = 0.065\pm0.005$. Here, we expressed the ellipticity and shear in quasi-cartesian coordinates, rather than polar coordinates, as the distribution of uncertainties is more gaussian. The two coordinate systems are related by $(e_c,\gamma_c) = (e,\gamma)\cos2(\theta_e,\theta_\gamma)$ and $(e_s,\gamma_s) = (e,\gamma)\sin2(\theta_e,\theta_\gamma)$. The derived image magnifications are approximately 19, 11, 11, and 0.7 for A, B, C and D, respectively. The uncertainties were determined using MCMC as plotted in Fig.~\ref{mcmc_gravlens}, where the convergence of the chains was again checked using the method of \citet{gel95}. The $\chi^2$ for this model is 0.8.

While the parameters derived above are independent of redshift, estimating the expected time delays between the images does require redshift information. Using the results of Section 3, we have fixed the redshift of the source at 0.57. If we assume a source redshift of 0.82, according to the best-fit result of the \citet{wu10} method in Section 3, we derive from the expression of the Einstein radius in a singular isothermal profile, $b=4\pi\left(\frac{\sigma}{c}\right)^2\frac{D_{ls}}{D_{s}}$, a velocity dispersion of $\sigma\sim500$ km/s. Here $c$ is the speed of light, and $D_{s},D_{ls}$ are the angular diameter distances to the source and between lens and source, respectively. This value is too large for a single galaxy, but if we employ the second-best source redshift estimate, $z\sim2.6$, which is preferred by LePhare, the combination of WISE and PS1 colors (Section 3) and the variability results in Section 4, we obtain $\sigma\sim315$ km/s, a more common value for massive early-type galaxies. We therefore assume this redshift to infer the predicted time delays, and we obtain $\Delta t_{BC} = 0.1$ days, $\Delta t_{AC} = 1.7$ days and $\Delta t_{DC} = 226.2$ days, with image C (a minimum of the time delay surface) leading. Image B is another minimum, and images A and D are saddle points.

@@

We would like to thank Arunav Kundu, Shobita Satyapal, Nathan Secrest, Bryan Dorland and Valeri Makarov for useful discussions and help. C.E.R was funded through the NSF grant AST-1312329, “Collaborative Research: Accurate cosmology with strong gravitational lens time delays”, and the HST grant GO-12889.
The Pan-STARRS1 Surveys (PS1) and the PS1 public science archive have been made possible through contributions by the Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, the Max-Planck Society and its participating institutes, the Max Planck Institute for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, the Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las Cumbres Observatory Global Telescope Network Incorporated, the National Central University of Taiwan, the Space Telescope Science Institute, the National Aeronautics and Space Administration under Grant No. NNX08AR22G issued through the Planetary Science Division of the NASA Science Mission Directorate, the National Science Foundation Grant No. AST-1238877, the University of Maryland, Eotvos Lorand University (ELTE), the Los Alamos National Laboratory, and the Gordon and Betty Moore Foundation.
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.

@@

\bibitem[McKean et al.(2007)]{mck07} McKean J.~P., et al., 2007, MNRAS, 378, 109 
\bibitem[Moffat (1969)]{mof69} Moffat, A. F. J.\ 1969, \aap, 3, 455
\bibitem[More et al.(2016)]{more16} More A., et al., 2016, MNRAS, 456, 1595
\bibitem[Myers et al.(2003)]{mye03} Myers S.~T., et al., 2003, MNRAS, 341, 1

@@

\begin{figure}
\epsscale{1.0}
\plotone{f3.eps}
\caption{
Covariances between the parameters used to model the observed system in $i$-band with Hostlens. The three contour lines mark the 1-, 2- and 3-$\sigma$ limits. The parameters are, in order: FWHM in arcseconds, ellipticity, orientation, shape of the first Moffat profile (1,2,3,4), same for the second Moffat profile (5,6,7,8), relative flux as flux1/(flux1 + flux2) (9), sky level in counts (10), positions in pixels on the x and y axes of point-source A (11,12) and its flux in counts (13), and same for objects B,C,D (14-25).}
\label{mcmc_hostlens}
\end{figure}
